---
title: "Gold_price"
author: "Abulmohsen Alhafith"
date: "18\x0F/2\x0F/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Loading Packages
```{r }
library(tidyverse)     # for data manipulation
library(rio)           # for importing data
library(caTools)       # for splittin the data 
```
## Impoting Data
```{r importing data}
gold_price<- import("data/gold.csv")


```

## Exploring the Data 
```{r}
summary(gold_price)


head(gold_price)
tail(gold_price)
```
This data has the prices of gold in US Dollar from January 1950 to June 2020 and published by https://datahub.io/core/gold-prices .

## Sitting up the Data 

```{r change the date to numbers}
r<- c(1:length(gold_price$Date))     # a set of number represent the number of the months 


gold_price$Date <- r   # change date column  to number 


head(gold_price)         
```




## Splitting the Data 
```{r Splitting the data}
training <- gold_price[1:677, ]              # training set 
testing  <- gold_price[678:847, ]            #  testing set 


summary(training)
summary(testing)

```


## Correlation
```{r correlation}
cor(gold_price$Date,gold_price$Price)            # to find the correlation
```
 There is strong correlation between the predictor variable and the respone variable.

## Model the Training Set with Linear Regression
```{r model the training set with linear regression }
model<- lm(formula = Price ~ Date, data = training)         # Price is the dependent variable and Date is independent variable  


summary (model)

```

We get the intercept is -31.1 and the slope is 0.722. 
when date increase by 1 the price would increse by .722.

## Predicting
```{r Predicting }
modeltest<-predict(model, newdata = testing)


```


```{r form new data to compare the model to the acutal values }
t<-data.frame(testing$Date,testing$Price,modeltest)
```

##  Visualizing Training Set

```{r Visualizing training set }
ggplot()+
  geom_point(aes(training$Date , y=training$Price), colour='red' ) +   # to plot the acutal values of training set
  geom_line (aes(x=training$Date, y= predict(model, newdata = training )) , colour= 'navy')+  # to plot the model for training set
  
  ggtitle ('Price of gold over time for training set   ' )+
  xlab('time')+ ylab('price') 
  

  

```


## Visualizing Testing set 
```{r Visualizing testing set }
ggplot()+
  geom_point(aes(t$testing.Date , y=t$testing.Price), colour='red' ) +   # to plot the acutal values of testing set
  geom_line (aes(x=t$testing.Date, y= t$modeltest ) , colour= 'navy')+  # to plot the model form training set in testing set 
  
  ggtitle ('Price of gold over time for tseting set   ' )+
  xlab('time')+ ylab('price') 
  

  

```